---
title: "In Depth Introduction"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
vignette: >
  %\VignetteIndexEntry{In Depth Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache = FALSE}
library(mlrMBO)
set.seed(123)
knitr::opts_chunk$set(cache = TRUE, collapse = FALSE, results = 'hold')
knitr::knit_hooks$set(document = function(x){
  gsub("```\n*```r*\n*", "", x)
})
```

This Vignette is supposed to give you a more in-depth introduction to **mlrMBO** and the principles and different approaches of model-based optimization.

## Objective Function
We use the Branin function for this example.
```{r}
objfun1 = makeBraninFunction()
ggplot2::autoplot(objfun1, render.levels = TRUE, show.optimum = TRUE)
```


## Initial Design

The second argument of the `mbo()` function - `design` - is the initial design with default setting `NULL`.

An easy (and recommended) way to create an initial design is to use the `generateDesign()` function from the **ParamHelpers** package.
If no design is given (i.e. `design = NULL`) a *Maximin Latin Hypercube* `lhs::maximinLHS()` design is used with `n = 4 * getNumberOfParameters(objfun1)` points.
Other possibilities to generate designs are for example `generateGridDesign` and `generateRandomDesign`.

_Note:_ If special designs are desired (e.g., orthogonal designs), they can be given as a `data.frame` but you have to make sure that the output resembles the output of `generateDesign()`.

For our objective function `objfun1` we create a slightly larger number of initial points than the default suggests.

```{r}
n = 5 * getNumberOfParameters(objfun1)
des = generateDesign(n = n, par.set = getParamSet(objfun1), fun = lhs::randomLHS)
```

If the design does not include the results of the objective function *mbo* will automatically calculate them in advance.
Sometimes it makes sense to precalculate the results because you might want to reuse the design.
In our case it is quite easy:

```{r}
des$y = apply(des, 1, objfun1)
```

_Note:_ *mlrMBO* uses `y` as a default name for the outcome of the objective function.
This can be changed in the control object.


## Surrogate Model

In our example we use *Kiriging* as a surrogate because it is the most common for numerical model-based optimization and has proven to be quite effective.
We use **mlr** to generate the Kriging regression learner from the package **DiceKriging**.

```{r}
surr.km = makeLearner("regr.km", predict.type = "se", covtype = "matern3_2", control = list(trace = FALSE))
```

In fact you can use any other regression learner from **mlr** as a surrogate.
Depending on the *infill criterion* we will set later it is important that the learner can predict the *uncertainty* (defined by `predict.type = "se"`) alongside the *mean* prediction.
Another popular surrogate would be the random forest.
It's use is explained in the Vignette for *mixed space optimization*.


## MBOControl

The `MBOControl` object controls the optimization process and is created with `makeMBOControl()`.
All parameters are documented in `?makeMBOControl`.
Some that influence the way MBO works are e.g.:
* `n.objectives`: The number of objectives. 1 implies normal single criteria optimization and is covered in this page. For values >1 see the Vignette for *multi-objective optimization*.
* `propose.points`: The number of evaluated points in each iteration. The default is 1 and refers to the standard SMBO process. Higher values suggest that you want to do *parallelization* (see Vignette).
* `final.method`: Defines how the final solution is proposed from the finished optimization process. The default is to return the best reached value of `y` and its corresponding `x`. However, this can give misleading results for noisy functions.


### MBOControlInfill

With `setMBOControlInfill()` you can change the default infill criterion settings in the `MBOControl` object.
It is highly recommended to adjust the settings to suit your optimization problem and the surrogate model.

#### Argument _crit_

An important role plays the choice of how the next design points in the sequential loop are chosen.
5 different infill criteria can be set via the `crit` argument in `setMBOControlInfill()`:

* `mean`: mean response of the surrogate model.
* `ei`: expected improvement of the surrogate model, which is the recommended setting if you use a *Kriging* surrogate.
* `aei`: augmented expected improvement, which is especially useful for noisy functions.
* `eqi`: expected quantile improvement.
* `cb`: confidence bound, which is the additive combination of mean response and mean standard error estimation of the surrogate model ($\hat{\mu} - \lambda \cdot \hat{s}$).

Here you can also further configure the infill criterion (e.g. `crit.cb.lambds` for the lambda parameter if `crit == "cb"`).

_Note:_ When using *Kriging* as a surrogate, numerical problems can occur if training points are too close to each other.
To circumvent this problem you can set `filter.proposed.points` to `TRUE`.
Then points closer then the value of `filter.proposed.points.tol` to an already evaluated point will be replaced with a random point.


#### Argument _opt_

The key idea behind model-based optimization is to substitute the expensive optimization on the black-box with optimization on the surrogate as this is deemed to be cheaper.
To optimize the infill criterion on the surrogate we also need an optimizer.
The optimum of the infill criterion function gives us the candidate to be evaluated on the expensive function.
Which one to use can be defined with the `opt` argument.
The possibilities are:

* `focussearch`: A Latin Hypercube design of size `opt.focussearch.points` (default 10000) is sampled in the parameter space (by `randomLHS`) and the design point with the best prediction of the infill criterion is determined. Then, the parameter space is shrunk around this point. This step is repeated `opt.focussearch.maxit` (default 5) times and the best observed point is passed back.
* `cmaes`: The optimal point is found with a covariance matrix adapting evolutionary strategy from the **cmaes** package. If the strategy fails, a random point is generated and a warning is given. Further control arguments can be provided in `opt.cmaes.control` as a list.
* `ea`: Use an evolutionary multi-objective optimization algorithm from the package **emoa** to determine the best point. The population size mu can be set by  `opt.ea.mu` (default value is 10). (mu+1) means that in each population only one child is generated using crossover and mutation operators. The parameters `eta` and `p` of the latter two operators can be adjusted via the attributes `opt.ea.sbx.eta`, `opt.ea.sbx.p`,`opt.ea.pm.eta` and `opt.ea.pm.p`. The default number of EA iterations is 500 and can be changed by `opt.ea.maxit` attribute.
* `nsga2`: Use the non-dominated sorting genetic algorithm from the package **nsga2R** to determine the best point. This algorithm should be used for *multi objective optimization* (see Vignette).

As all four infill optimization strategies do not guarantee to find the global optimum, users can set the number of restarts by the `opt.restarts` argument (default value is 1).
Of all restarts the best infill criterion value determines the candidate to be evaluated.

_Note:_ Only the `focussearch` optimizer is suitable for for categorical parameters in the parameter set.


### setMBOControlTermination

With `setMBOControlTermination()` you can set multiple stopping criteria and the first one that is met will terminate the optimization process, e.g:

* `iters`: The maximum number of iterations
* `time.budget`: A maximum running time in seconds
* `target.fun.value`: A threshold for function evaluation (stop if a evaluation is better than a given value)
* `max.evals`: The maximum number of function evaluations

_Note:_ You can also create your own stopping condition(s).

### setMBOControlMultiPoint

This extends a MBO control object with options for multi-point proposal.
Multi-point proposal means, that multiple points are proposed and evaluated, which is especially useful for parallel batch evaluation. For a detailed introduction, check the Vignette on parallelization.

### setMBOControlMultiCrit

This adds multi-criteria optimization specific options to the control object. For details see the Vignette on *multi-criteria optimization*.


### Experiments and Output

Although we presented various options we will mainly go with the defaults settings.

```{r eval=TRUE}
ctrl = makeMBOControl()
ctrl = setMBOControlInfill(
  control = ctrl,
  crit = makeMBOInfillCritEI()
)
ctrl = setMBOControlTermination(
  control = ctrl,
  iters = 10
)
```

Now we will apply the `mbo()` function to optimize the objective function.

```{r, eval=TRUE}
res = mbo(objfun1, design = des, learner = surr.km, control = ctrl, show.info = FALSE)
print(res)
```

The default output of mbo contains the best found parameter set and the optimization path.
The `MBOResult` object contains additional information, most importantly:

* `x`: The best point of the parameter space
* `y`: The associated best value of the objective function
* `opt.path`: The optimization path. See `?ParamHelpers::OptPath` for further information.
* `models`: Depending on `store.model.at` in the `MBOControl` object, this contains zero, one or multiple surrogate models (default is to save the model generated at the last iteration).
* ... (see `?MBOSingleObjResult`)

We can also change some arguments of the `MBOControl` object and run `mbo()` again:

```{r, eval=FALSE}
ctrl = setMBOControlInfill(ctrl, crit = makeMBOInfillCriterCB())
ctrl = setMBOControlTermination(ctrl, iters = 5L)
mbo(objfun1, design = des, learner = surr.km, control = ctrl, show.info = FALSE)
```
